---
title: "Incarceration of Asian American in the U.S. from 1990 to 2016"
author: "Jimmy Cheong"
date: "2/23/2022"
output: html_document
theme: ""
---

```{r setup, include=FALSE}
library(dplyr)
library(dbplyr)
library(ggplot2)
library(tidyverse)
library(maps)
library(knitr)
source("../source/analysis.R")

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(out.width='1000px', dpi=300)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Introduction
This project analysis incarceration trends in the U.S. to learn the situations/trends of incarceration of Asian American by comparing its trend to trend of incarceration of Whites.
This report focuses on the count and rate of **Prison Population**, **Jail Population** of each races and compare their trends(*Prison/Jail Rate*, *Prison/Jail Rate Change*) against each other to learn the difference of incarceration trend by race.

To be specific, we would put our eyes on variables including:

`Population Count (Ages 15 to 64)`, `Jail Population Count`, and `Prison Population Count`

among *AAPI(Asian American/Pacific Islander)* and *Whites*.

However, this report only analysis data since 1990 to 2016 as data outside this period 1990 is mostly missing, incomplete, and were partly estimated.

### Summary Information

```{r summary,echo=FALSE}
kable(summary_table)

# Average Prison population per year: `r summary$average_prison_pop`
# Highest Prison population in a year: `r summary$highest_p_rate_in_a_year`
# Lowest Prison population in a year: `r summary$lowest_p_rate_in_a_year`
# Highest Jail population in a year: `r summary$highest_j_rate_in_a_year`
# Lowest Jail population in a year: `r summary$lowest_j_rate_in_a_year`
# Prison increasement since 1990 in percentage: `r summary$p_increasement`
# Jail increasement since 1990 in percentage: `r summary$j_increasement`
```


\

# Trends over time chart
The following chart shows the percentage population of AAPI/WHITE in Prison/Jail. This shows the trend in population rate who got into prison and jail for every year from 1990 to 2016.

We could see that while both Prison/Jail population rate had increased over time for both races. The increasement of White is steeper while increasement of AAPI is has a relatively gentle slope.

For the design, I've picked a color blind friendly color pattern and matched two sets of colors to each of the race*(Orange for AAPI, Blue for White)*.
```{r tsc,echo=FALSE}
plot(tsc)
```

\

# Variable comparison chart

The following two charts had shown and compared two trends involving percentage increasement of prison population and percentage increasement of total population from 1990 to 2016.
As shown in the left chart, the population increasement of White in the period is very gentle while the population of AAPI had increase over 250% in the period.

We should expect the percentage prison population of AAPI also increase in a similar fashion compare to AAPI total population's increasement. Just as the percentage jail population of AAPI does.
However, the percentage increasement prison population of AAPI is no where near to **250%** but far greater, reaching over **4000%** since the year 2000.
Which is 40 times the prison population back in 1990. On the other hand, the Prison/Jail population increase of White only **increased a little bit**.

This surely shows a pattern where the rate of AAPI arrested into prison is greatly higher than White do.

:::: {style="display: flex;"}

::: {}

```{r Variable comparison chart1,echo=FALSE}
plot(chart_population_increasement)
```
:::

::: {}

```{r Variable comparison chart2,echo=FALSE}
plot(chart_prison_jail_rate_increasement)
```
:::

::::

\

# Map

This map had shown the difference in average prison rate of AAPI and White for each county.

On The map was drawn by counties to help us spot specific counties which most bias on AAPI prison incarceration.

The more **red** the county shape, the **higher** the AAPI prison rate compare to White, vice versa.

We can see that there are a few county with absolute **red** *(biased toward White with higher AAPI prison rate)* on the east side of the U.S. and an absolute **no red** county *(biased toward AAPI with higher White prison rate)* near the center part.

There're also a few state lack of data and marked as gray on the map.
```{r map,echo=FALSE}
plot(map)
```
